---
title: "Central Limit Theorem"
author: "Rohit Tolawat"
date: "6/12/2020"
output: html_document
---

```{r loadpackages, warning=FALSE, message=FALSE}
pacman::p_load(caret, data.table, ggplot2, dplyr)
knitr::opts_chunk$set(echo = FALSE, fig.width=12, fig.height=6, fig.path = 'Figs/')
options(digits = 3)
```


#Reading the hospital dataset
```{r Reading the .csv file from home directory}
hospitalData <- read.csv("C:/Users/rohit/Downloads/diabetic_data.csv", header = TRUE)
dim(hospitalData)
```

```{r}
head(hospitalData)
```

#Isolating the time in hospital column for better understanding and manipulation
```{r}
hospitalData.dt <- setDT(hospitalData)
#Checking if there are any NA rows in the column
sum(is.na(hospitalData.dt$time_in_hospital))
sqrt(var(hospitalData.dt$time_in_hospital))
```

#Checking the distribution of the population and the mean days spent by a patient in hospital
```{r}
timeDist <-  ggplot(hospitalData.dt, aes(x=hospitalData.dt$time_in_hospital)) +
              geom_histogram(color = "black", fill = "white", binwidth = 1) 
timeDist + geom_vline(xintercept =mean(hospitalData.dt$time_in_hospital), color = "red", linetype = "dashed")   + 
          ggtitle("Distribution of days in hospital(Dashed line = Population mean)") +
          xlab("Days in hospital") + ylab("Number of patients")

```

```{r}
mean(hospitalData.dt$time_in_hospital)
sqrt(var(hospitalData.dt$time_in_hospital))
```


#Insight corner: Most patients are discharged in less than 5 days, the distributed is right-skewed with an average of 4.4 days per patient

#Taking 1000 samples of size 2,10,20 and 40 from the population
```{r}
set.seed(123)
sampleMeanSize2 <- c()
sampleMeanSize10 <- c()
sampleMeanSize20 <- c()
sampleMeanSize40 <- c()
numberOfSamples = 1000

for(i in 1:numberOfSamples){
sampleMeanSize2[i] <- mean(sample(hospitalData.dt$time_in_hospital,size = 2, replace = TRUE))
sampleMeanSize10[i] <- mean(sample(hospitalData.dt$time_in_hospital,size = 10, replace = TRUE))
sampleMeanSize20[i] <- mean(sample(hospitalData.dt$time_in_hospital,size = 20, replace = TRUE))
sampleMeanSize40[i] <- mean(sample(hospitalData.dt$time_in_hospital,size = 40, replace = TRUE))
}
par(mfrow=c(2,2))
hist(sampleMeanSize2,main = "1000 samples of 2", xlab = "Days in hospital" )
abline(v=mean(sampleMeanSize2), col = "red")
hist(sampleMeanSize10,main = "1000 samples of 10", xlab = "Days in hospital" )
abline(v=mean(sampleMeanSize10), col = "red")
hist(sampleMeanSize20,main = "1000 samples of 20", xlab = "Days in hospital" )
abline(v=mean(sampleMeanSize20), col = "red")
hist(sampleMeanSize40,main = "1000 samples of 40", xlab = "Days in hospital" )
abline(v=mean(sampleMeanSize40), col = "red")

```
#The Q-Q plot can be used to check if the data points form a normal distribution

```{r}
qqnorm(sampleMeanSize40)
```


```{r}
Mu <- mean(sampleMeanSize40)
err <- sqrt(var(sampleMeanSize40)) / sqrt(40)
sigmaBar <- sqrt(var(sampleMeanSize40))
sigmaPop <- sqrt(var(sampleMeanSize40))*sqrt(40)
cumulative.prob <- 0
daysInHospital <- 0
j <- 0
for (i in seq(0,20,0.001)){
  if(i == min(sampleMeanSize40))
    j = 1  
  else
    j = j + 1
  
  cumulative.prob[j] <- pnorm(i, mean = Mu,  sd = sigmaPop, lower.tail = TRUE)
  daysInHospital[j] <-  i 
}
cumulative.dt <- data.table(var1 = cumulative.prob*100, var2 = daysInHospital)

ggplot(cumulative.dt, aes(y=var1, x=var2))+
  geom_line()+
  xlab("Days in hospital")+
  ylab("Proportion of patients treated in %")+
  ggtitle("Cumulative probability distribution plot from population parameters")
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")

```

#Insight corner: Sampling distribution of sample means attains normal distriburion as sample size increases 
```{r}
sample_size <- c(2,10,20,40)
sample_mean <- c(mean(sampleMeanSize2),mean(sampleMeanSize10),mean(sampleMeanSize20),mean(sampleMeanSize40))
sample_sd   <- c(sd(sampleMeanSize2) ,sd(sampleMeanSize10),sd(sampleMeanSize20),sd(sampleMeanSize40))
result.dt <- data.table(sample_size,sample_mean,sample_sd)
names(result.dt) <- c("Sample size", "Sample mean", "Sample SD")
result.dt
```

#Plotting standard deviation against sample size
```{r}
ggplot(result.dt, aes(x=result.dt$`Sample size`, y=result.dt$`Sample SD`)) +
              geom_line() +
              geom_point(size = 3)+
ggtitle("Variation in the estimate decreases as sample size increases") +
xlab("Sample size") + ylab("Standard deviation")

```

#Plotting estimate of the mean against sample size
```{r}
ggplot(result.dt, aes(x=result.dt$`Sample size`, y=result.dt$`Sample mean`)) +
              geom_line() +
              geom_point(size = 3)+
geom_hline(yintercept =mean(hospitalData.dt$time_in_hospital),color = "red", linetype = "dashed")   +    
ggtitle("As sample size increases, estimate inches closer to population mean") +
xlab("Sample size") + ylab("Estimate mean of population")

```
```{r}
controlChartSampleMean <- 0
controlChartSD <- 0

twoSigmaBar = 2*sigmaBar
threeSigmaBar = 3*sigmaBar 
  
for(i in 1:40){
  sampleDrawn <- sample(hospitalData.dt$time_in_hospital,size = 10, replace = TRUE)
  controlChartSampleMean[i] <- mean(sampleDrawn)
  controlChartSD[i] <- sqrt(var(sampleDrawn))
}

controlChartData.dt <- data.table(var1 = 1:40, var2=controlChartSampleMean, var3=controlChartSD)

MuBarBar <- mean(controlChartSampleMean)

plot1 <- ggplot(controlChartData.dt, aes(y=controlChartData.dt$var2, x=controlChartData.dt$var1))+
  geom_line(size=0.5)+
  geom_point(size=3)+
  geom_hline(yintercept = MuBarBar, linetype = "dashed", color = "Blue", size=0.5)+
  geom_hline(yintercept = (MuBarBar + sigmaBar), linetype = "dashed", color = "green",size=0.5)+
  geom_hline(yintercept = (MuBarBar - sigmaBar), linetype = "dashed", color = "green",size=0.5)+
  geom_hline(yintercept = (MuBarBar + twoSigmaBar), linetype = "dashed", color = "yellow",size=0.5)+
  geom_hline(yintercept = (MuBarBar - twoSigmaBar), linetype = "dashed", color = "yellow",size=0.5)+  
  geom_hline(yintercept = (MuBarBar + threeSigmaBar), linetype = "dashed", color = "red",size=0.5)+
  geom_hline(yintercept = (MuBarBar - threeSigmaBar), linetype = "dashed", color = "red",size=0.5)+  
  xlab("Subgroup sample number")+
  ylab("Subgroup sample mean")+
  ggtitle("Monitoring chart for subgroup sample mean")

plot2 <- ggplot(controlChartData.dt, aes(x=controlChartData.dt$var1, y=controlChartData.dt$var3))+
 geom_line()+
 geom_point(size=3)+
 geom_hline(yintercept = mean(controlChartData.dt$var3), linetype="dashed", color = "Blue", size=1)+
 geom_hline(yintercept = 1.716 * mean(controlChartData.dt$var3), linetype="dashed", color = "red", size=0.5)+
 geom_hline(yintercept = 0.284 * mean(controlChartData.dt$var3), linetype="dashed", color = "red", size=0.5)+
 ggtitle("Monitoring Chart for subgroup standard deviation")+
 xlab("Subgroup sample number")+
 ylab("Standard deviation")  
  
plot1
plot2

     
```


```{r}

seg1 <- sum(controlChartData.dt[,var2>=(MuBarBar-sigmaBar) & var2<=(MuBarBar+sigmaBar)])
seg2 <- sum(controlChartData.dt[,var2>=(MuBarBar-twoSigmaBar) & var2<=(MuBarBar+twoSigmaBar)])
seg3 <- sum(controlChartData.dt[,var2>=(MuBarBar-threeSigmaBar) & var2<=(MuBarBar+threeSigmaBar)])
seg4 <- nrow(controlChartData.dt) - seg3

segmentName <- c("Within 1 sigma", "Within 2 Sigma", "Within 3 Sigma", "Beyond Control Limits") 
controlTable.dt <- data.table(var1=segmentName, var2=c(seg1,seg2,seg3,seg4), var3=c(seg1/40*100,seg2/40*100,seg3/40*100,seg4/40*100))
names(controlTable.dt) <- c("Sigma limits","Subgroup mean within limits","Percentage points within limits")
controlTable.dt
```

```{r}
mean(controlChartData.dt$var2)
mean(controlChartData.dt$var3)
ttest <-  t.test(controlChartData.dt$var2, mu = Mu)

ttest
ttest$p.value

```



