---
title: "Central Limit Theorem"
author: "Rohit Tolawat"
date: "6/12/2020"
output: html_document
---

```{r loadpackages, warning=FALSE, message=FALSE}
pacman::p_load(caret, data.table, ggplot2, dplyr)
knitr::opts_chunk$set(echo = FALSE, fig.width=12, fig.height=6, fig.path = 'Figs/')
options(digits = 3)
```


#Reading the hospital dataset
```{r Reading the .csv file from home directory}
hospitalData <- read.csv("C:/Users/rohit/Downloads/diabetic_data.csv", header = TRUE)
dim(hospitalData)
```

```{r}
head(hospitalData)
```

#Isolating the time in hospital column for better understanding and manipulation
```{r}
hospitalData.dt <- setDT(hospitalData)
#Checking if there are any NA rows in the column
sum(is.na(hospitalData.dt$time_in_hospital))
```

#Checking the distribution of the population and the mean days spent by a patient in hospital
```{r}
timeDist <-  ggplot(hospitalData.dt, aes(x=hospitalData.dt$time_in_hospital)) +
              geom_histogram(color = "black", fill = "white", binwidth = 1) 
timeDist + geom_vline(hospitalData.dt, xintercept =mean(hospitalData.dt$time_in_hospital), color = "red", linetype = "dashed")   + 
          ggtitle("Distribution of days in hospital(Dashed line = Population mean)") +
          xlab("Days in hospital") + ylab("Number of patients")

```
#Insight corner: Most patients are discharged in less than 5 days, the distributed is right-skewed with an average of 4.4 days per patient

#Taking 1000 samples of size 2,10,20 and 40 from the population
```{r}
set.seed(123)
sampleMeanSize2 <- c()
sampleMeanSize10 <- c()
sampleMeanSize20 <- c()
sampleMeanSize40 <- c()
numberOfSamples = 1000

for(i in 1:numberOfSamples){
sampleMeanSize2[i] <- mean(sample(hospitalData.dt$time_in_hospital,size = 2, replace = TRUE))
sampleMeanSize10[i] <- mean(sample(hospitalData.dt$time_in_hospital,size = 10, replace = TRUE))
sampleMeanSize20[i] <- mean(sample(hospitalData.dt$time_in_hospital,size = 20, replace = TRUE))
sampleMeanSize40[i] <- mean(sample(hospitalData.dt$time_in_hospital,size = 40, replace = TRUE))
}
par(mfrow=c(2,2))
hist(sampleMeanSize2,main = "1000 samples of 2", xlab = "Days in hospital" )
abline(v=mean(sampleMeanSize2), col = "red")
hist(sampleMeanSize10,main = "1000 samples of 10", xlab = "Days in hospital" )
abline(v=mean(sampleMeanSize10), col = "red")
hist(sampleMeanSize20,main = "1000 samples of 20", xlab = "Days in hospital" )
abline(v=mean(sampleMeanSize20), col = "red")
hist(sampleMeanSize40,main = "1000 samples of 40", xlab = "Days in hospital" )
abline(v=mean(sampleMeanSize40), col = "red")
```
#Insight corner: Sampling distribution of sample means attains normal distriburion as sample size increases 
```{r}
sample_size <- c(2,10,20,40)
sample_mean <- c(mean(sampleMeanSize2),mean(sampleMeanSize10),mean(sampleMeanSize20),mean(sampleMeanSize40))
sample_sd   <- c(sd(sampleMeanSize2),sd(sampleMeanSize10),sd(sampleMeanSize20),sd(sampleMeanSize40))
result.dt <- data.table(sample_size,sample_mean,sample_sd)
result.dt
```

#Plotting standard deviation against sample size
```{r}
ggplot(result.dt, aes(x=result.dt$sample_size, y=result.dt$sample_sd)) +
              geom_line() +
ggtitle("Variation in the estimate decreases as sample size increases") +
xlab("Sample size") + ylab("Standard deviation")

```

#Plotting estimate of the mean against sample size
```{r}
ggplot(result.dt, aes(x=result.dt$sample_size, y=result.dt$sample_mean)) +
              geom_line() +
geom_hline(hospitalData.dt, yintercept =mean(hospitalData.dt$time_in_hospital),color = "red", linetype = "dashed")   +    
ggtitle("As sample size increases, estimate inches closer to population mean") +
xlab("Sample size") + ylab("Estimate mean of population")

```

